Limit Theory for Out-of-Sample Comparisons of Overfit Models

### 2012-03-30 paper.tex
Collapsed hold history into a single commit

\documentclass[11pt]{article}
%\linespread{1.2}\selectfont
\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{fit}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[round,comma]{natbib}
\usepackage{setspace,amsmath,amssymb,amsthm,ifthen,subfig,url,multicol}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\thepage}\chead{}\rhead{}
\lfoot{}\cfoot{}\rfoot{}
\renewcommand\headrulewidth{0pt}
\onehalfspacing
\usepackage[top=1.222222in,left=0.9444444in,right=1.888889in,bottom=2.4444444in]{geometry}
\usepackage{nath}

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{claim}[thm]{Claim}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{asmp}{Assumption}
\newtheorem{example}{Example}
\newtheorem{defn}{Definition}

\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator*{\plim}{plim}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\atan}{atan*}

@@

\newcommand{\oost}{\ensuremath{\operatorname{oos-t}}}
\newcommand{\oosSum}[2]{\ensuremath{\sum_{#1=R-\h+#2}^{T-\h}}}

@@

\newcommand{\ZSummand}{D_{sR} - \E_R D_{sR}}
\newcommand{\ZDef}{ \frac1{\sqrt{P}} \SumInnerBlock{s}{i} [\ZSummand]}
\newcommand{\ZSqCE}{\ensuremath{\CEOuterBlock{Z_i^2}{i}}}
\newcommand{\ZTrunc}{\ensuremath{C\sqrt{b_{T}/P}}}
\newcommand{\PScaleTerm}{\ensuremath{\frac{C \sqrt{b_T}}{P}}}

\newcommand{\vWeight}{W((t-s)/\gamma)}
\newcommand{\vSummand}{\ensuremath{(D_{tR} - \E_R D_{tR})(D_{sR} - \E_R D_{sR}) \vWeight}}
\newcommand{\kernelBDefn}[1]{\ensuremath{\frac1{\delta\sqrt{2\pi}}e^{-\frac{#1^2}{2\delta^2}}}}
\newcommand{\kernelB}[1]{\ensuremath{\eta_{\delta}(#1)}}

@@

\newcommand{\varianceTermIDefn}{\ensuremath{\frac1{P}\oosSum{s,t}{1}
    \vSummand}}
\newcommand{\varianceTermI}{\ensuremath{\sigma_{0\delta}^2}}
\newcommand{\varianceTermII}{\ensuremath{\sigma_{1\delta}^2}}
\newcommand{\varianceTermIIa}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{l=\max(1-t,-P)}^{\min(P-t,P)} (D_{t+l+R,R} - \E_R D_{t+l+R,R})
    W(l/\gamma)}}
\newcommand{\varianceTermIIb}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{j=1-t}^{P-t} (D_{t+j+R,R} - \E_R D_{t+j+R,R})
    \kernelB{j/\gamma}}}
\newcommand{\varianceTermIII}{\ensuremath{\sigma_{2\delta}^{2}}}
\newcommand{\varianceTermIIIa}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{l=\vttLower}^{\vttUpper} (D_{t+l+R,R} - \E_R D_{t+l+R,R})
    W(l/\gamma)}}
\newcommand{\varianceTermIV}{\ensuremath{\sigma_{3\delta}^{2}}}
\newcommand{\varianceTermIVb}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{j=\vttLower}^{\vttUpper} (D_{t+j+R,R} - \E_R D_{t+j+R,R})
    \kernelB{j/\gamma}}}
\newcommand{\vtIIIsummand}{\ensuremath{(Z_{1t}Z_{2t} - \E_RZ_{1t}Z_{2t})}}

\newcommand{\varianceDiffA}{\ensuremath{\frac1{P} \oosSum{s,t}{1}
    \lvert (D_{tR} - \E_R D_{tR})(\E_R D_{sR} - \bar D_1) \rvert
    \vWeight}}
\newcommand{\varianceDiffAi}{\ensuremath{\frac1P \oosSum{s}{1}(\E_R
    D_{sR} - \bar D_1 )^2}}
\newcommand{\varianceDiffAii}{\ensuremath{\frac1P \oosSum{s}{1}
    (D_{sR} - \E_R D_{sR})^2}}

@@

\newcommand{\dmw}{\textsc{dmw}}
\newcommand{\ols}{\textsc{ols}}
\newcommand{\mds}{\textsc{mds}}
\newcommand{\clt}{\textsc{clt}}
\newcommand{\lln}{\textsc{lln}}
\newacronym{VAR}{Vector Autoregression}{VAR}
\newacronym{oos}{out-of-sample}{OOS}
\newcommand{\dgp}{\textsc{dgp}}
\newcommand{\hac}{\textsc{hac}}
\newacronym{ned}{Near Epoch Dependent}{NED}
\newacronym{mse}{Mean Squared Error}{MSE}
\newcommand{\sfield}{$\sigma$-field}
\newcommand{\sfields}{\sfield s}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}

\frenchspacing \title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I
    would like to thank Julian Betts, Marjorie Flavin, Nir Jaimovich,
    Lutz Kilian, Michael McCracken, Seth Pruitt, Ross Starr, Yixiao
    Sun, Allan Timmermann, participants at the 2010 Midwest Economics
    Association Annual Meetings, the 2010 Joint Statistical Meetings,
    and in many seminars at \textsc{ucsd}, and espeically Graham
    Elliott for their valuable suggestions, feedback and advice in
    writing this paper.  I would also like to thank Amit Goyal for
    providing computer code and data for his 2008 RFS paper with Ivo
    Welch \citep{goyal-welch-2008-rfs}.} \\ Iowa State University,
  Economics Department}
\date{October 24, 2011}

\begin{document}
\maketitle

\begin{abstract}\thispagestyle{empty}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the number of
  predictors used by the larger model increase with the number of
  observations so that their ratio remains uniformly positive.  Under
  this limit theory, the naive Diebold-Mariano-West out-of-sample test
  can test hypotheses about a key quantity for evaluating forecasting
  models---a time series analogue to the generalization error---as
  long as the out-of-sample period is small relative to the total
  sample size.  Moreover, tests that are designed to reject if the
  larger model is true, such as the usual in-sample Wald and
  \textsc{lm} tests and also Clark and McCracken's (2001, 2005a),
  \citepos{mccracken_asymptotics_2007} and Clark and West's (2006,
  2007) out-of-sample statistics, will choose the larger model too
  often when the smaller model is more accurate.

@@

\noindent Keywords: Generalization Error, Forecasting, Model
Selection, t-test, Dimension Asymptotics
\end{abstract}

@@

Although \oos\ comparisons have been popular in macroeconomics and
finance since \citepos{meese_empirical_1983} seminal study of exchange
rate models, it is unclear from a theoretical perspective whether or
not the statistics are useful.  Empirical researchers often cite
``overfit'' or ``instability'' as reasons for using \oos\ comparisons,
as in \citet{stock_forecasting_2003}, but neither term is precisely
defined or formalized.  Compounding this problem, the asymptotic
distributions of these statistics are derived under conditions that
rule out either instability or overfit and allow a researcher to use a
conventional in-sample comparison---a variation of the F-test, for
example.  As \citet{inoue_kilian_2004} argue, the statistics
themselves are designed to test hypotheses that can be tested by these
in-sample statistics.  For example, \citet{diebold_comparing_1995} and
\citet{west_asymptotic_1996} derive the limiting distributions of many
popular \oos\ test statistics under conditions that would justify
these full-sample tests.\footnote{In this paper, we will refer to
  their basic \oos\ $t$-test as the \dmw\ test.}  Much of the
subsequent research by \citet{chao_out_2001},
\citet{clark_tests_2001,clark_evaluating_2005},
\citet{corradi_consistent_2002,corradi_recent_2004},
\citet{clark_using_2006,clark_approximately_2007},
\citet{mccracken_asymptotics_2007}, \citet{anatolyev_inference_2008},
and others relaxes several of Diebold and Mariano's and West's
assumptions, but maintains the stationarity and dependence conditions
that permit in-sample comparisons.\footnote{Like us,
  \citet{anatolyev_inference_2008} allows the number of regressors to
  increase with $T$.  But in that paper, they increase slowly enough
  that the \ols\ coefficients are consistent and asymptotically
  normal.}  \citet{giacomini_tests_2006}, and
\citet{giacomini-rossi-2009-REStud,giacomini-rossi-2010-JAppliedEconometrics}
are exceptions.  Instead of focusing on hypotheses that can be tested
by in-sample comparisons, \citet{giacomini_tests_2006} derive an \oos\
test for the null hypothesis that the difference between two models'
\oos\ forecasting performance is unpredictable, a martingale
difference sequence, \citet{giacomini-rossi-2009-REStud} test whether
the \oos\ forecasting performance of a model suffers from a breakdown
relative to its in-sample performance, and
\citet{giacomini-rossi-2010-JAppliedEconometrics} test whether the
forecasting performance is stable. However, those papers focus on a
particular \oos\ estimation strategy and do not address why naive
\oos\ comparisons might be generally useful.

Since in-sample and \oos\ statistics require similar assumptions and
test similar hypotheses, one might expect that they would give similar
results.  They do not.  Generally in-sample analysis supports more
complicated theoretical models and \oos\ analysis supports simple
benchmarks, as seen in \citet{meese_empirical_1983},
\citet{stock_forecasting_2003}, and \citet{goyal-welch-2008-rfs}.
Since these different approaches strongly influence the outcome of
research, it is important to know when each is appropriate.  The
explanations in favor of \oos\ comparisons claim that they should be
more robust to unmodeled instability
\citep{clark_power_2005,giacomini_tests_2006,giacomini-rossi-2009-REStud,giacomini-rossi-2010-JAppliedEconometrics}
or to overfit \citep{mccracken_data_1998,clark_can_2004}.  Both
explanations presume that the in-sample comparison is invalid and the
\oos\ comparisons are more reliable.  Of course, as
\citet{inoue_kilian_2004,inoue_selection_2006} point out, both
in-sample and \oos\ methods could be valid, but the \oos\ methods
could have lower power.

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of \oos\ comparisons when at least
one of the models is overfit---when the number of its regressors
increases with the number of observations so that their ratio remains
positive.  We focus on linear regression models estimated with a fixed
training window, but our basic conclusions should be true for other
estimation strategies and models as well.  We show that the usual
\oos\ average is asymptotically normal and can consistently estimate
the difference between the models' expected loss in periods
$T+1,T+2,\dots$ conditional on the on the data available in period
$T$, a time-series analogue to the generalization error.\footnote{See,
  for example, \citet{elemstatlearning}.}  Under these asymptotics,
the generalization error does not converge to the expected performance
of the pseudo-true models, so the in-sample and \oos\
comparisons measure different quantities and should be expected to
give different results for reasons beyond simple size and power
comparisons.  Under our theory, the model that is closer to the true
\dgp\ in population can forecast worse.  In such a situation, an
in-sample comparison would correctly reject the null hypothesis that
the benchmark is true, and an \oos\ comparison would correctly fail to
reject the null that the benchmark is more accurate.\footnote{In a
  pair of papers similar to ours,
  \citet{RePEc:fip:fedkrw:rwp09-10,RePEc:fip:fedkrw:rwp09-11} study
  in-sample and \oos\ tests that the larger model has nonzero
  coefficients that are too close to zero to expect the model to
  forecast more accurately.  Like this paper, they argue that the
  larger model can be true but less accurate.  However, they focus on
  an aspect of the \dgp\ that makes this phenomenon likely, while we
  focus on the coefficient estimates that produce less accurate
  forecasts.}

This result justifies \oos\ comparisons when researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in these settings, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.\footnote{These costs can be
  monetary, in the form of computer upgrades, software installation,
  and training, or they can be reputational costs.}
We show that the \dmw\ test lets the forecaster control the
probability of the first error, just as with conventional hypothesis
testing.

However, since the models' coefficients are imprecisely
estimated, the test sample must be small to ensure that the model
estimated over the training sample is similar to the one that will be
estimated over the full sample.  In particular, $P/T \to 0$ is
required for consistent estimation, and $P^2/T \to 0$ is required for
valid confidence intervals and inference.  For larger $P$, the \oos\
comparisons remain asymptotically normal, but are centered on the
forecasting performance associated with the period-$R$ estimates.  In
practice, researchers typically use large values of $P$, so these
studies may be too pessimistic about their models' future accuracy if
they use the \dmw\ test.  Section~\ref{sec:oostheory} lays out the
asymptotic behavior of the \dmw\ test under this limit theory.

Under this asymptotic theory, the \ols\ coefficient estimator is no
longer consistent or asymptotically normal \citep{huber_robust_1973},
so this is a useful way to model overfit and imprecisely estimated
models.  Moreover, \citet{efron_biased_1986,efron_estimation_2004}
shows that in-sample comparisons overestimate models' accuracy by a
factor proportional to this ratio.  Despite the non-normality of the
coefficient estimators, modifications of the F-test are valid under
this asymptotic theory, as shown by \citet{boos_anova_1995},
\citet{akritas_asymptotics_2000},
\citet{akritas_heteroscedastic_2004}, and
\citet{calhoun_hypothesis_2009}, among others.  The F-test and Wald
tests are no longer asymptotically valid themselves.  Such results
establish that hypotheses about the parameters of the \dgp\ can be
tested in-sample even if one or both models are overfit.  For that
very reason, these tests can not reliably choose a model for
forecasting and will reject the benchmark too often.  We show this
analytically in Section~\ref{sec:insample}.  Moreover, under this
asymptotic theory, many recent \oos\ test statistics, such as those
derived by \cite{clark_tests_2001,clark_evaluating_2005},
\citet{mccracken_asymptotics_2007}, and
\citet{clark_using_2006,clark_approximately_2007} should behave like
the F-test.  These tests are also designed to reject the benchmark
when the alternative model is true, and so they may reject too often
when the benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
\infty$ (with the number of regressors fixed), these statistics behave
like the \dmw\ test when $P$ is small, but should over-reject the
benchmark when $P$ is large.  Section~\ref{sec:mc} presents
simulations that support these claims.

Finally, this paper also introduces a new method of proof for \oos\
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of \oos\ loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales, so asymptotic results for \ned\
functions of mixing processes can be used for these \oos\ processes
with only slight modification.  Although we only present a proof that
this mixingale result holds for the fixed-window scheme, it is simple
to extend it to other estimation windows.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
\dmw\ \oos\ test and shows that the F-test rejects the benchmark model
too often when it is false but more accurate than the
alternative. Section~\ref{sec:mc} presents a Monte Carlo study
supporting our theoretical results.  Section~\ref{sec:empirics}
applies the \oos\ statistic to \citepos{goyal-welch-2008-rfs} dataset
for equity premium prediction, and Section~\ref{sec:conclusion}
concludes.  Proofs and supporting results are listed in the Appendix.

>>>>

\documentclass[11pt]{article}
% \makeatletter
% \renewcommand\@seccntformat[1]{\protect\makebox[0pt][r]{\csname
%     the#1\endcsname\quad}}
% \renewcommand\section{\@startsection
% {section}{1}{0mm}{\baselineskip}{\baselineskip}{\normalfont\normalsize\scshape}}
% \renewcommand\subsection{\@startsection
% {subsection}{2}{0mm}{\baselineskip}{\baselineskip}{\normalfont\normalsize\scshape}}
% \renewcommand\subsubsection{\@startsection
% {subsubsection}{3}{0mm}{\baselineskip}{\baselineskip}{\normalfont\normalsize\scshape}}
% \renewcommand\paragraph{\@startsection
% {paragraph}{4}{0mm}{\baselineskip}{\baselineskip}{\normalfont\normalsize\scshape}}
% \renewcommand\subparagraph{\@startsection
% {subparagraph}{5}{0mm}{\baselineskip}{\baselineskip}{\normalfont\normalsize\scshape}}
% \makeatother
\linespread{1.2}\selectfont
\usepackage{tikz}\usetikzlibrary{calc}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[round,comma]{natbib}
\usepackage{setspace,amsmath,amssymb,amsthm,ifthen,subfig}
\usepackage[top=1in,left=1in,right=1in,bottom=1in]{geometry}
%\onehalfspacing
%\usepackage[top=1.222222in,left=0.9444444in,right=1.888889in,bottom=2.4444444in]{geometry}
\usepackage{nath}
% \newcommand{\wall}{\,}
% \newcommand{\return}{\,}

%\newtheoremstyle{gc}{\topsep}{\topsep}{\itshape}{0pt}{\scshape}{}{\newline}{}
%\theoremstyle{gc}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{claim}[thm]{Claim}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{asmp}{Assumption}
\newtheorem{example}{Example}
\newtheorem{defn}{Definition}

% \newcommand{\E}{E}
% \newcommand{\var}{var}
% \newcommand{\plim}{plim}
% \DeclareMathOperator*{\argmin}{arg\,min}
% \newcommand{\tr}{tr}
\newcommand{\E}{`E}
\newcommand{\var}{`var}
\newcommand{\plim}{`plim}
\newcommand{\tr}{`tr}

@@

% \newcommand{\oost}{\ensuremath{\operatorname{oos-t}}}
\newcommand{\oost}{\ensuremath{'oos-'t}}
\newcommand{\oosSum}[2]{\ensuremath{\sum_{#1=R+#2}^{T-\h}}}

@@

\newcommand{\ZSummand}{D_{sR} - \E_R D_{sR}}
\newcommand{\ZDef}{ \frac1{\sqrt{P}} \SumInnerBlock{s}{i} [\ZSummand]}
\newcommand{\ZSqCE}{\ensuremath{\CEOuterBlock{Z_i^2}{i}}}
\newcommand{\ZTrunc}{\ensuremath{C\sqrt{\frac{b_{T}}{P}}}}
\newcommand{\PScaleTerm}{\ensuremath{\frac{C \sqrt{b_T}}{P}}}

\newcommand{\vWeight}{W(\frac{t-s}{\gamma})}
\newcommand{\vSummand}{\ensuremath{(D_{tR} - \E_R D_{tR})(D_{sR} - \E_R D_{sR}) \vWeight}}
\newcommand{\kernelBDefn}[1]{\ensuremath{\frac1{\delta\sqrt{2\pi}}e^{-\frac{#1^2}{2\delta^2}}}}
\newcommand{\kernelB}[1]{\ensuremath{\tau_{\delta}(#1)}}

@@

\newcommand{\varianceTermIDefn}{\ensuremath{\frac1{P}\oosSum{s,t}{1}
    \vSummand}}
\newcommand{\varianceTermI}{\ensuremath{\sigma_{0\delta}^2}}
\newcommand{\varianceTermII}{\ensuremath{\sigma_{1\delta}^2}}
\newcommand{\varianceTermIIa}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{l=\max(1-t,-P)}^{\min(P-t,P)} (D_{t+l+R,R} - \E_R D_{t+l+R,R})
    W(\frac{l}{\gamma})}}
\newcommand{\varianceTermIIb}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{j=1-t}^{P-t} (D_{t+j+R,R} - \E_R D_{t+j+R,R})
    \kernelB{\frac{j}{\gamma}}}}
\newcommand{\varianceTermIII}{\ensuremath{\sigma_{2\delta}^{2}}}
\newcommand{\varianceTermIIIa}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{l=\vttLower}^{\vttUpper} (D_{t+l+R,R} - \E_R D_{t+l+R,R})
    W(\frac{l}{\gamma})}}
\newcommand{\varianceTermIV}{\ensuremath{\sigma_{3\delta}^{2}}}
\newcommand{\varianceTermIVb}{\ensuremath{\frac1{\sqrt{P\gamma}}
    \sum_{j=\vttLower}^{\vttUpper} (D_{t+j+R,R} - \E_R D_{t+j+R,R})
    \kernelB{\frac{j}{\gamma}}}}
\newcommand{\vtIIIsummand}{\ensuremath{(Z_{1t}Z_{2t} - \E_RZ_{1t}Z_{2t})}}

\newcommand{\varianceDiffA}{\ensuremath{\frac1{P} \oosSum{s,t}{1}
    \lvert (D_{tR} - \E_R D_{tR})(\E_R D_{sR} - \bar D_1) \rvert
\vWeight}}
\newcommand{\varianceDiffAi}{\ensuremath{\frac1P \oosSum{s}{1}(\bar
    D_1 - \E_R D_{sR})^2}}
\newcommand{\varianceDiffAii}{\ensuremath{\frac1P \oosSum{s}{1}
    (D_{sR} - \E_R D_{sR})^2}}

@@

\newacronym{ols}{Ordinary Least Squares}{OLS}
\newacronym{mds}{Martingale Difference Sequence}{MDS}
\newacronym{clt}{Central Limit Theorem}{CLT}
\newacronym{lln}{Law of Large Numbers}{LLN}
\newacronym{VAR}{Vector Autoregression}{VAR}
\newacronym{oos}{out-of-sample}{OOS}
\newacronym{dgp}{Data Generating Process}{DGP}
\newacronym{hac}{Heteroskedasticty- and Autocorrelation-Consistent}{HAC}
\newacronym{ned}{Near Epoch Dependent}{NED}
\newacronym{mse}{Mean Squared Error}{MSE}
\newcommand{\sfield}{$\sigma$-field}
\newcommand{\sfields}{\sfield s}
\renewcommand{\Re}{\ensuremath{\mathbb{R}}}

\frenchspacing 
\title{Limit Theory for Out-of-Sample Comparisons of Overfit Models} 
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I would like to thank
    Julian Betts, Robert de Jong, Marjorie Flavin, Nir Jaimovich,
    Michael McCracken, Seth Pruitt, Ross Starr, Yixiao Sun, Allan
    Timmermann, participants at the Midwest Economics Association
    Annual Meetings and in many seminars at \textsc{ucsd}, and espeically
    Graham Elliott for their valuable suggestions, feedback and advice in writing
    this paper.  I would also like to thank Amit Goyal for providing
    computer code and revised data for his 2008 RFS paper with Ivo
    Welch \citep{goyal-welch-2008-rfs}.} \\ Iowa State University, Economics Department}

\begin{document}
\bibliographystyle{abbrvnat}
\maketitle

\begin{abstract}\noindent
  This paper uses dimension asymptotics to find a reason to compare
  overfit forecasts out-of-sample instead of in-sample. The two models
  to be compared are linear regressions, and the number of predictions
  used by the larger model increases with the number of observations
  so that their ratio remains uniformly positive. Under this limit
  theory, tests that are designed to reject if the larger model is
  true, such as the usual in-sample Wald and \textsc{lm} tests and
  also Clark and McCracken's (2001, 2005a),
  \citepos{mccracken_asymptotics_2007} and Clark and West's (2006,
  2007) out-of-sample statistics, will choose the larger model too
  often when the smaller model is more accurate. I show that the
  Gaussian out-of-sample $t$-test is a valid test that the models have
  equal generalization error as long as the out-of-sample period is
  very small relative to the total sample size.

@@

\noindent Keywords: Generalization Error, Forecasting, Model
Selection, t-test
\end{abstract}

@@

Although \oos\ comparisons have been popular in macroeconomics and
finance since \citepos{meese_empirical_1983} seminal study of exchange
rate models, it is unclear from a theoretical perspective whether or
not the statistics are useful.  Empirical researchers often cite {\em
  overfit} or {\em instability} as reasons for using \oos\
comparisons, as in \citet{stock_forecasting_2003}, but neither term is
precisely defined or formalized.  Compounding this problem, the
asymptotic distributions of these statistics are derived under
conditions that rule out either instability or overfit and allow a
researcher to use a conventional in-sample comparison---a variation of
the F-test, for example.  The statistics themselves are designed to
test hypotheses that can be tested by these in-sample statistics.  For
example, \citet{diebold_comparing_1995} and
\citet{west_asymptotic_1996} derive the limiting distributions of many
popular \oos\ test statistics under conditions that would justify
these full-sample tests.
Much of the subsequent research by \citet{chao_out_2001},
\citet{clark_tests_2001,clark_evaluating_2005},
\citet{corradi_consistent_2002,corradi_recent_2004},
\citet{clark_using_2006,clark_approximately_2007}, and
\citet{mccracken_asymptotics_2007}, among others, relaxes several of
\citepos{diebold_comparing_1995} and \citepos{west_asymptotic_1996}
assumptions, but maintains the stationarity and dependence conditions
that permit in-sample comparisons.  \citepos{giacomini_tests_2006}
paper is an exception.  Instead of
focusing on hypotheses that can be tested by in-sample comparisons,
\citet{giacomini_tests_2006} derive an \oos\ test for the null
hypothesis that the difference between two models' \oos\ forecasting
performance is unpredictable. However, \citet{giacomini_tests_2006}
focus on a particular \oos\ estimation strategy and do not address why
\oos\ comparisons might be useful in general.

Since in-sample and \oos\ statistics require similar assumptions and
test similar hypotheses, one might expect that they would agree in
practice.  They do not.  Generally in-sample analyses support more
complicated theoretical models and \oos\ analyses support simple
benchmarks, as seen in \citet{meese_empirical_1983} and
\citet{stock_forecasting_2003}.  Since these different approaches
strongly influence the outcome of research, it is important to know
when each is appropriate.  The explanations in favor of \oos\
comparisons claim that they should be more robust to unmodeled
instability \citep{clark_power_2005,giacomini_tests_2006} or to
overfit
\citep{mccracken_data_1998,clark_can_2004,hansen_in-sample_2009}.
Both explanations presume that the in-sample comparison is invalid and
the \oos\ comparisons are more reliable.  Of course, as
\citet{inoue_in-sample_2005,inoue_selection_2006} propose, both
in-sample and \oos\ methods could be valid, but the \oos\ methods
could have lower power.

This paper uses dimension asymptotics to study the behavior of \oos\
comparisons when at least one of the models is overfit---the number of
regressors increases with the number of observations so that their
ratio remains positive.  We focus on linear regression models
estimated with a fixed training window, but our basic conclusions
should be true for other schemes as well.  We show that the usual
\oos\ average is asymptotically normal and consistently estimates the
difference between the models' generalization error, the expected loss
of the estimated models' forecasts for new observations.  Under our
asymptotics, the generalization error does not converge to the
expected performance of the pseudo-true models.  Therefore, the
in-sample and \oos\ comparisons measure different quantities and
should be expected to give different results for reasons beyond simple
size and power comparisons.  Under our theory, the model that is
closer to the true \dgp\ in population can forecast worse.  In such a
situation, an in-sample comparison would correctly reject the null
hypothesis that the benchmark is true, and an \oos\ comparison would
correctly fail to reject the null that the benchmark is more
accurate.\footnote{In a pair of papers similar to ours,
  \citet{RePEc:fip:fedkrw:rwp09-10,RePEc:fip:fedkrw:rwp09-11} study
  in-sample and \oos\ tests that the larger model has nonzero
  coefficients that are too close to zero to expect the model to
  forecast more accurately.  Like this paper, they argue that the
  larger model can be true but less accurate.  However, they focus on
  an aspect of the \dgp\ that makes this phenomenon likely, while we
  focus on the coefficient estimates that produce less accurate
  forecasts.}

Under this asymptotic theory, the \ols\ coefficient estimator is no longer
consistent or asymptotically normal \citep{huber_robust_1973}, so this
is a constructive way to model overfit and imprecisely estimated
models.  Moreover, \citet{efron_biased_1986,efron_estimation_2004}
shows that in-sample comparisons overestimate models' accuracy by a
factor proportional to this ratio.  Despite the non-normality of the
coefficient estimators, modifications of the F-test are valid
under this asymptotic theory, as shown by \citet{boos_anova_1995},
\citet{akritas_asymptotics_2000},
\citet{akritas_heteroscedastic_2004}, and
\citet{calhoun_hypothesis_2009}, among others.  The F-test and Wald
tests are no longer asymptotically valid themselves.

Moreover, under this asymptotic theory, non-standard test statistics,
such as those derived by \citet{mccracken_asymptotics_2007} and
\cite{clark_tests_2001,clark_evaluating_2005}, are unnecessary, as in
\citet{giacomini_tests_2006}.  Gaussian confidence intervals
constructed over the test sample contain the difference between the
models' future performance with the correct probability (i.e. 95\%
confidence intervals contain it 95\% of the time).  In fact, using
these non-Gaussian test statistics, or correcting the \oos\ average as
in \citet{clark_using_2006,clark_approximately_2007}, will cause the
forecaster to choose the larger model too often.  Since these tests
are designed to reject if the benchmark model is false, they are
invalid tests for equal generalization error in exactly the same way
that the in-sample tests are: they can overreject when the larger model
is true but less accurate.  Since these corrections vanish as $P/T$
converges to zero, this criticism does not hold for small
test-samples.

However, there is a trade off.  Since we allow the coefficients to be
imprecisely estimated, the test sample must be small to ensure that
the model estimated over the training sample is similar to the one
that will be estimated over the full sample.  In particular, $P/T \to
0$ is required for the \oos\ average to consistently estimate the
generalization error, and $P^2/T \to 0$ is required for valid
confidence intervals.  For larger $P$, the \oos\ comparisons remain
asymptotically normal, but are centered on the generalization error
associated with the wrong estimates.
In practice, researchers typically use large values of $P$, so these
studies may be too pessimistic about their models' future accuracy.

Our asymptotic theory is similar to that developed by
\citet{anatolyev_inference_2008}, but there are several key
distinctions.  \citet{anatolyev_inference_2008} derives the asymptotic
distribution of \oos\ comparisons as the number of regressors
increases with the number of observations, but slowly: their ratio
does not remain positive in the limit.  Anatolyev's asymptotics
motivate several corrections to the \oos\ statistic for it to continue
to give valid inference about the models' population parameters; those
corrections are unnecessary under our asymptotics because we focus on
inference about the models' generalization error, not their
pseudo-true parameters, and they would (likely) be insufficient to
give valid inference about the population parameters under our faster
growth rates for the number of regressors.

This paper also introduces a new method of proof for \oos\ processes.
We use a coupling argument (Berbee's 1979 Lemma) to show that these
processes behave like Mixingales when the underlying series are
absolutely regular, even if the forecasts depend on full-sample
non-convergent estimators.  Moreover, transformations of these
processes also behave like Mixingales, so asymptotic results for \ned\
functions of mixing processes can be used on these \oos\ processes
with slight modification.  Although we only present a proof that this
mixingale result holds for the fixed-window scheme, it is simple to
extend it to other estimation windows.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces this paper's notation and
assumptions.  Section~\ref{sec:theory} gives the main theoretical
results.  Section~\ref{sec:empirics} applies the \oos\ statistic to
the \citet{goyal-welch-2008-rfs} dataset.  Section~\ref{sec:mc}
presents a Monte Carlo study, and section~\ref{sec:conclusion}
concludes.  Proofs and additional results are listed in the Appendix.

\section{Setup and Assumptions}\label{sec:assumptions}
We will setup our models and notation first, then list the assumptions
required for our theory.  There are two competing linear models that
give forecasts for the target, $y_{t+\h}$:
\[
y_{t+\h} = x_{1t}'\theta_1 + \varepsilon_{1,t+h}
\]
and
\[
y_{t+\h} = x_{2t}'\theta_2 + \varepsilon_{2,t+h};
\]
$\h$ is the forecast horizon, and the variables $y_t$, $x_{1t}$, and
$x_{2t}$ are all known in period $t$.  Let
\begin{equation*}
  \mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)
\end{equation*}
with $x_t$ the vector of all elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$.  We will allow $K_1$ and $K_2$
to vary with $T$, so a stochastic array underlies all of our
asymptotic theory, but we suppress that notation to make the
presentation simpler.

Although our theory can apply to many different \oos\ statistics,
we'll only present results for the \oos\ t-test (\oost).  Both models
are estimated using \ols\ over the training sample:
\[
\hat y_{i,t+\h} = x_{it}'\bh{iR}, \quad i=1,2, \quad t = R+1,\dots,T-h
\]
with
\[
\bh{it} = (\sum_{s=1}^{t-\h} x_{is}x_{is}')^{-1} \sum_{s=1}^{t-\h} x_{is} y_{s+\h},
\quad i=1,2.
\]
The \oos\ average is denoted $\bar D_1$:
\[
\bar D_1 \equiv P^{-1} \oosSum{t}{1} D_{tR}
\quad \text{with} \quad D_{tS} \equiv L(y_{t+\h} - x_{1t}'\bh{1S}) -
L(y_{t+\h} - x_{2t}'\bh{2S}),
\]
$P = T - R$ and $L$ a known loss function.  The \oos\ statistic
of interest is
\[
\oost \equiv \sqrt{P} \frac{\bar D_1}{\hat\sigma_1},
\]
for $\hat\sigma^2$ some (possibly \hac) estimator of the asymptotic
variance of $\bar D_1$.

We'll show that \oost\ can be used to test that the models have equal
generalization error.  To motivate interest in this
quantity, assume that whichever model is chosen will be used for an
additional $Q$ periods in the future.  Let $\bar D_2$ denote the
difference in average performance over those periods:
\[
\bar D_2 \equiv Q^{-1} \sum_{t=T+1}^{T+Q} D_{tT}.
\]
Note that $\bar D_1$ uses the training sample estimators,
$\hat\theta_{iR}$, and $\bar D_2$ uses the full sample estimators,
$\hat\theta_{iT}$.  The forecaster has access to the information set
$\mathcal{F}_T$ when deciding between the two forecasting models, but
$\mathcal{F}_T$ does not contain $\bar D_2$.  So the forecaster should
decide between them based on the conditional expectation, $\E_T
\bar{D}_2$. If this quantity is positive, the first model is expected
to perform worse in the future, and if it is negative the second is.
